DCMIP2016: A Review of Non-hydrostatic Dynamical Core Design and Intercomparison of Participating Models

### 2017-08-13 DCMIP2016-Part1-EditorResponse.tex
Updates based on returned reviews

\noindent Atmospheric dynamical cores are a fundamental component of global atmospheric modeling systems, and are responsible for capturing the dynamical behavior of the Earth's atmosphere via numerical integration of the Navi\'er-Stokes equations.  These systems have existed in one form or another for over half of a century, with the earliest discretizations having now evolved into a complex ecosystem of algorithms and computational strategies.  In essence, no two dynamical cores are alike, and their individual successes suggest that no perfect model exists.  To better understand modern dynamical cores, this paper aims to provide a comprehensive review of eleven dynamical cores, drawn from modeling centers and groups that participated in the 2016 Dynamical Core Model Intercomparison Project (DCMIP) workshop and summer school.  This review includes choice of model grid, variable placement, vertical coordinate, prognostic equations, temporal discretization, and the diffusion, stabilization, filters and fixers employed by each system.

>>>>

\noindent \textbf{Response to Editor} \\

\noindent A throughly revised code availability table has been included with the latest version of the paper.  Six of the modeling groups have uploaded model source code to Zenodo.  Licensing agreements prevent the remainder from directly posting model code publicly, however nearly all provide an avenue for public access to the codebase.  Only two of the model groups (ACME-A and FVM) cannot allow public access at this time, but plan for a public release of the code in the future.  In the mean time, these groups will allow the topical editor (Julia Hargreaves) access the code.  The following page contains the updated code availability table.

### 2017-08-13 DCMIP2016-Part1-Models-20170604.tex
Updates based on returned reviews

@@

%GEM is a semi-Lagrangian model with a hybrid coordinate of a log-hydrostatic-pressure type, here denoted $\zeta$, described in section \ref{sec:GEM_zeta}.
In GEM the non-hydrostatic equations are written explicitly as deviations from hydrostatic balance represented by
\begin{align}
\mu = \pdiff{p}{s} - 1,
\end{align}
\noindent where $s$ (denoted $\pi$ in GEM documentation) is given by (\ref{eq:s_mass-based}).  In this case the equations of GEM model \citep{Girard2014} are concisely given by
\begin{align}
\frac{d \vb{u}_h }{dt} + f {\bf k} \times {\bf{u}_{h}} + R_d T_v\nabla_\zeta \,  \log\, p+ \left( 1 + \mu \right) \nabla_\zeta {\Phi} &= 0, \\
\frac{d w}{dt}- g_c \mu &= 0, \label{eq:GEMVerticalVelocity} \\
\frac{d}{dt} \log \left( \pdiff{s}{\zeta}\right) + \nabla_{ \zeta }\cdot {\bf u}_{{h}}+\frac{\partial \dot{ \zeta }}{\partial \zeta } &= 0, \\
\frac{d \log T_v}{dt} - \frac{R_d}{c_p} \frac {d \log p} {dt} &= 0, \\
\frac{\partial \Phi}{\partial s} + \frac{R_dT_v}{p} &= 0, \\
\frac{d \Phi}{dt}-  g_c w & = 0.
\end{align}  Here $\nabla_\zeta$ denotes the horizontal gradient along $\zeta$ surfaces.

@@

>>>>

 \usepackage{xcolor} \usepackage{ulem} \usepackage{changebar}
 \newcommand\TLSins[1]{\cbstart{}\textcolor{ins}{\uline{#1}}\cbend{}}
 \newcommand\TLSdel[1]{\cbdelete{}\textcolor{del}{\sout{#1}}}

@@

%GEM is a semi-Lagrangian model with a hybrid coordinate of a log-hydrostatic-pressure type, here denoted $\zeta$, described in section \ref{sec:GEM_zeta}.
In GEM the non-hydrostatic equations are written explicitly as deviations from hydrostatic balance represented by
\begin{align}
\mu = \pdiff{p}{s} - 1,
\end{align}
\noindent where $s$ (denoted $\pi$ in GEM documentation) is given by (\ref{eq:s_mass-based}).  In this case the equations of GEM model \citep{Girard2014} are concisely given by
\begin{align}
\frac{d \vb{u}_h }{dt} + f {\bf k} \times {\bf{u}_{h}} + R_d T_v\nabla_\zeta \,  \log\, p+ \left( 1 + \mu \right) \nabla_\zeta {\Phi} &= 0, \\
\frac{d w}{dt}- g_c \mu &= 0, \label{eq:GEMVerticalVelocity} \\
\frac{d}{dt} \log \left( \pdiff{s}{\zeta}\right) + \nabla_{ \zeta }\cdot {\bf u}_{{h}}+\frac{\partial \dot{ \zeta }}{\partial \zeta } &= 0, \\
\frac{d \log T_v}{dt} - \frac{R_d}{c_p} \frac {d \log p} {dt} &= 0, \\
\frac{\partial \Phi}{\partial s} + \frac{R_dT_v}{p} &= 0, \\
\frac{d \Phi}{dt}-  g_c w & = 0.
\end{align}  Here $\nabla_\zeta$ denotes the horizontal gradient along $\zeta$ surfaces.  With respect to the treatment of moisture in GEM, the cloud water and all non-gases are embedded in the total air density $\rho$,
affecting the virtual temperature defined in (\ref{eq:VirtualTemperatureExact}). Also, specific mass is used in GEM (not mixing ratio).

@@

Information on the availability of source code for the models featured in this paper is tabulated below.

\begin{center}
\begin{tabular}{cp{5in}}
\hline Short Name & Code availability \\ \hline 
ACME--A & ACME, including ACME--A, is under active development funded by the U.S. Department of Energy.  ACME version 1.0 is scheduled to be publicly released under an open source license in 2018, but is not available at present.$^\dagger$ \\
CSU & CSU model source code is available under the BSD 3-clause license.  The release used for DCMIP2016 can be found via Zenodo (\mbox{\url{http://dx.doi.org/10.5281/zenodo.580099}}). \\
DYNAMICO & DYNAMICO is open source and available online from IPSL Forge (\mbox{\url{http://forge.ipsl.jussieu.fr/dynamico}}) or directly by request to Thomas Dubos (\mbox{dubos@lmd.polytechnique.fr}).  The release used for DCMIP2016 can be found via Zenodo (\mbox{\url{http://dx.doi.org/10.5281/zenodo.583718}}). \\
FV$^3$ & FV$^3$ model source code is available through the GFDL Virtual Lab (\url{https://vlab.ncep.noaa.gov/web/fv3gfs}).  Access requires users to create a Virtual Lab account. \\
FVM & Model codes developed at ECMWF, including the IFS and FVM, are intellectual 
property of ECMWF and its member states.  Although the FVM code is not publicly available at present, it is expected that FVM will be available in the near future under the OpenIFS license (\url{http://www.ecmwf.int/en/research/projects/openifs}).  The repo tag for the version of FVM that applies for DCMIP is ``v0.1''.$^\dagger$ \\
GEM & Due to licensing requirements, GEM model code is only available by request to Abdessamad Qaddouri (Abdessamad.Qaddouri@canada.ca) or Vivian Lee (Vivian.Lee2@canada.ca). \\
ICON & ICON is freely available to the scientific community for non-commercial research under an institutional license issued by project partners DWD+MPI-M.  Because of the restrictions of this license, access to the code is only available by request to G\"unther Z\"{a}ngl (\mbox{Guenther.zaengl@dwd.de}) or Marco Giorgetta (\mbox{marco.giorgetta@mpimet.mpg.de}). \\
\hline \multicolumn{2}{p{6in}}{$\dagger$ \footnotesize In compliance with the GMD editorial requirements, this code has been made available to the topical editor in charge of the manuscript.}
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{cp{5in}}
\hline Short Name & Code availability (cont'd) \\ \hline 
MPAS & Open-source via GitHub (\url{https://github.com/MPAS-Dev/MPAS-Release}).  The release used for DCMIP2016 can be found via Zenodo (\mbox{\url{http://dx.doi.org/10.5281/zenodo.583316}}) \\
NICAM & NICAM source code is available under the BSD 2-clause license via Zenodo (\mbox{\url{http://dx.doi.org/10.5281/zenodo.580128}}).  Further information on collaborating with the NICAM team can be found at \mbox{\url{http://nicam.jp/hiki/?Research+Collaborations}}. \\
OLAM & OLAM is open source and available online via SourceForge (\url{https://sourceforge.net/projects/olam-model/}).  The release used for DCMIP2016 can be found via Zenodo (\mbox{\url{http://doi.org/10.5281/zenodo.582308}}). \\
TEMPEST & Tempest source code is available under the Lesser GNU Public License on GitHub (\mbox{\url{https://github.com/paullric/tempestmodel}}).  The release used for DCMIP2016 can be found via Zenodo (\mbox{\url{http://dx.doi.org/10.5281/zenodo.579649}}). \\
\hline 
\end{tabular}
\end{center}

\clearpage
